preciseTAD: A machine learning framework for precise TAD boundary prediction at base pair resolution

### 2020-07-11 Manuscript/manuscript.Rmd
A couple of thoughts

Several methods have been proposed to identify the genomic coordinates of TAD boundaries throughout the linear genome (reviewed in [@Zufferey:2018aa; @forcato2017comparison]), less so for chromatin loops [@rao20143d; @Ay:2014aa; @Heinz:2010aa].  However, a key limitation among most of these TAD- and loop callers, referred hereafter as TAD callers for brevity, is that they are heavily reliant on Hi-C data *resolution*. Resolution refers to the sizes of genomic bins used to segment the linear genome as a way to deal with low signal-to-noise incurred by the high dimensional nature of Hi-C data [@li2019comparison]. Lower resolutions contribute to larger bin sizes, which lead to uncertainty in the location of TAD boundaries. Due to high sequencing costs, most available Hi-C datasets have relatively low resolution, ranging from 25 kb to 100 kb. Because increasing the resolution of Hi-C data requires a quadratic increase in the total sequencing depth, obtaining high resolution remains difficult [@schmitt2016genome]. While these low resolution Hi-C datasets can be used to probe higher order genomic patterns such as A/B compartments, more refined resolutions are necessary to accurately define TADs and chromatin loops [@fudenberg2016formation].

@@

The performance of *preciseTAD* was found to be empirically better [??? As judged by what?] than both ARROWHEAD and Peakachu ground truth boundaries. Guided by both low-resolution Hi-C data and high-resolution genome annotation data, *preciseTAD* predicts base-level resolution boundaries, alleviating resolution limitations of Hi-C data. We show that, unlike ARROWHEAD, *preciseTAD* does not inflate the number of predicted boundaries [??? Where did you show this? TBD], instead, providing only the most biologically meaningful boundaries that demarcate regions of high inter-chromosomal interactions. *preciseTAD* boundaries were found to be more enriched for known architectural transcription factors including CTCF, RAD21, SMC3, and ZNF143 [??? So what?]. Likewise *preciseTAD* boundaries were found to be more conserved between GM12878 and K562 cell lines, further highlighting their biological significance. [??? So what?]

>>>>

Several methods have been proposed to identify the genomic coordinates of TAD boundaries throughout the linear genome (reviewed in [@Zufferey:2018aa; @forcato2017comparison]), less so for chromatin loops [@rao20143d; @Ay:2014aa; @Heinz:2010aa; @salameh2019supervised].  However, a key limitation among most of these TAD- and loop callers, referred hereafter as TAD callers for brevity, is that they are heavily reliant on Hi-C data *resolution*. Resolution refers to the sizes of genomic bins used to segment the linear genome as a way to deal with low signal-to-noise incurred by the high dimensional nature of Hi-C data [@li2019comparison]. Lower resolutions contribute to larger bin sizes, which lead to uncertainty in the location of TAD boundaries. Due to high sequencing costs, most available Hi-C datasets have relatively low resolution, ranging from 25 kb to 100 kb. Because increasing the resolution of Hi-C data requires a quadratic increase in the total sequencing depth, obtaining high resolution remains difficult [@schmitt2016genome]. While these low resolution Hi-C datasets can be used to probe higher order genomic patterns such as A/B compartments, more refined resolutions are necessary to accurately define TADs and chromatin loops [@fudenberg2016formation].

@@

The performance of *preciseTAD* was found to be empirically better [??? As judged by what?] than both ARROWHEAD and Peakachu ground truth boundaries. Guided by both low-resolution Hi-C data and high-resolution genome annotation data, *preciseTAD* predicts base-level resolution boundaries, alleviating resolution limitations of Hi-C data. We show that, unlike ARROWHEAD, *preciseTAD* does not inflate the number of predicted boundaries [??? Where did you show this? TBD], instead, providing only the most biologically meaningful boundaries that demarcate regions of high inter-chromosomal interactions. *preciseTAD* TAD and loop boundaries were enriched for known architectural transcription factors including CTCF, RAD21, SMC3, and ZNF143, supporting recent observations that TADs and loops are dynamic structures generated by similar mechanisms [@hansen2018recent].  [??? So what?]. Likewise *preciseTAD* boundaries were found to be more conserved between GM12878 and K562 cell lines, further highlighting their biological significance. [??? So what?]
